### 2016-09-29 cdcl.tex
CDCL

\section{Lifting CDCL to numeric domains}
The work of~\cite{dhk2013-popl,sas12,tacas12} shows that 
satisfiability solvers are static analyzers. They present 
a framework that strictly generalizes CDCL algorithm to 
lattices and abstract transformers. We refer the reader 
to~\cite{dhk2013-popl} for a more formal perspective alongwith 
soundness and completeness proofs.  In this section, we focus 
on the practical aspect of lifting CDCL algorithm to numeric 
abstract domains.

We treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
$P$. 
CDCL algorithm has two main phases which feedback information to 
each other -- {\em model search} and {\em conflict analysis}. The model 
search procedure tries to find a satisfying assignment to the input 
formula. If the search fails, meaning that the current partial assignment 
leads to a {\em conflict}, the conflict analysis phase learns the reason 
for such failure to block the model search from entering into same 
conflicting search-space in future. More formally, given a set of 
formula $Form$ and a set of structure $Struct$, a semantic entailment 
relation, $\models \in \mathcal{P}(Struct \times Form)$, identifies the 
structure that satisfies a formula. A concrete domain is the power set 
of structures which gives us lattice. 
We define two transformers for a CDCL solver. \\
\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\

\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
that contradict a formula $\varphi$. This is formally shows as follows:
$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
\\
We now define abstract interpretation analogues to model search and conflict analysis.   \\
\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
set of traces generated by the program and is erroneous. This can be
characterised in two different ways. We can start with the initial state and
compute least fix-point with strongest post-condition or start from error state 
and compute the greatest fix-point with weakest pre-condition. \\

\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
set of program traces that are safe, that is, all program traces that are not
generated by the program. \\

\textit{Interaction between amod and aconf:} If the search for finding an 
erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
proof stage, we get back refinement of strongest post-condition which keeps in 
mind satisfiability information.