### 2016-10-12 related_work.tex
Related work

\section{Related Work}
The natural domain SMT framework proposed in~\cite{ndsmt} operates on
a fixed first-order partial assignment lattice structure, where first-order 
variables are mapped to domain values, similar to constants lattice 
in program analysis.  However, ACDCL can be instantiated on different 
abstract lattices.  The abstract lattice in natural domain SMT does 
not have complementable meet irreducibles, and therefore does not 
support generalized clause learning~\cite{sas}.  On the other hand, 
most abstract lattice have complementation property, thus enabling 
ACDCL to perform generalized clause learning.