### 2016-10-19 conclusion.tex

\section{Conclusions}
In this paper, we instantiate ACDCL solver with relational domains.  We explained specific properties of model search and conflict analysis procedure necessary
for such lifting.  Experimental evaluation on different benchmark sets shows that
ACDCL is able to suitably exploit the expressivity of richer
abstract domains as well as precision of CDCL architecture.
This is attributed to the intelligent decision heuristics
which exploits the high-level structure of the problem
combined with stronger deductions and clause learning
mechanism aided by the richer abstract domains.

\rmcmt{Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check. Implement dedicated abstract domains to exploit full performance.
generalization for relational domain, instantiate acdl with trace-based domains.}