### 2016-10-20 contributions.tex
example resolve

\paragraph{Contributions}
In this paper, we make the following contributions.
\begin{compactenum}
\item We present a novel program analysis technique that 
lifts the CDCL algorithm to relational numerical abstract domains.   
We have implemented our technique in 2LS verification tool.  
\Omit{
To the best of our knowledge, this is the first instantiation of CDCL 
algorithm with relational abstract domains. 
}

\item We present an abstract model search algorithm that strictly 
generalizes the Boolean Constraint Propagation (BCP) procedure to
compute a fixed point over a relational domain.  Our abstract 
conflict analysis algorithm present a non-trivial lifting of the 
Unique Implication Point (UIP) procedure to relational abstract 
domains in order to learn a generalized reason for conflicts.

\item  Experiments indicate two order of magnitude reduction 
in the number of decisions, propagations, and conflicts compared 
to a SAT solver while being more precise compared to an abstract interpreter.  
\Omit {
set of benchmarks drawn from the bitvector regression category in 
SV-COMP, bit-precise software models auto-generated from hardware 
circuits domain and some hand crafted example show that 
}
\end{compactenum}