### 2016-10-20 example.tex

\section{Motivating Example}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & Control-Flow Graph & Abstract Conflict Graph \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void foo(int v, 
         _Bool c) 
{
  assume(v>=0 $\wedge$ v<=N);
  int x; 
  if(c)
    x = v;
  else 
    x = -v;
  int z = x * x;
  assert(z>=0);
}
\end{lstlisting}
&
\begin{minipage}{3.7cm}
\centering
%\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{minipage}{4.50cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
\\
\hline
\end{tabular}
\caption{\label{fig:example}
C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
\end{figure}
Let us consider a simple program in Figure~\ref{fig:example} that performs 
multiplication of two machine integers and checks that the result of the 
multiplication is always positive for an integer input $v$.  To avoid 
overflow error, we assume $v$ has an upper bound $N$.  The control 
flow graph is shown in the middle column of Figure~\ref{fig:example}.

We analyze the program using three different analyzers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
\Omit {
The SAT solver makes 36 decisions, 1365 boolean constraint propagations 
and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
the solver performs a restart, presumably to recover from bad decisions.  
}
Standard BMC tools convert the program to a bit-vector equation and pass 
them to a SAT solver.  However, standard forward interval analysis is 
too imprecise to verify the safety of the above program.  The 
imprecision is due to control-flow join at node $n4$.  State-of-the-art 
abstract interpretation tool, Astr{\'e}e, requires external hints, provided 
by manually annotating the code with partition directives at $n1$, 
to prove safety.  The partition directive helps the tool to analyze 
each program execution paths separately.  To do so, Astr{\'e}e 
requires a product of trace partitioning domain and an interval domain 
to determine safety.  However, a relational domain such as octagons is 
able to prove the program without any partition.  
\Omit {
In general, the imprecision is either intended by the tool because such high 
precision analysis is normally not required for runtime error 
analysis or the imprecision is unavoidable due to the complexity 
of the application under analysis.  
}
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
\hline
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{solver}
\end{table}
%
The right-side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
octagon domain (top) and an interval domain (bottom).  The analysis 
associates an abstract element for each control location and variable, 
as shown in the abstract conflict graph in Firgure~\ref{fig:example}.  
Octagonal constraints deduced by ACDCL solver are sufficient to prove 
safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
an automatic interval analysis is also sufficient to prove safety with 
only a single decision.  Interval constraints generated from forward 
analysis in the initial deduction phase infers that 
$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.

Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
ACDCL implements several decision heuritics.  In this case, the 
branching variable $c$ is chosen by an {\em ordered} decision 
heuristics that select variables appearing in conditional branches 
first before choosing other program variables. 
The decision constrains the interval of $x:=[0,5]$.  The deductions made 
during fixed point iteration are represented by abstract conflict 
graph in right-side of Figure~\ref{fig:example}.  Nodes in the abstract 
conflict graph denotes the corresponding updates to the program 
variables in the CFG.  Interval analysis concludes that $(\text{Error}:\bot)$, that 
is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
the program is {\em safe} for $c:=[1,1]$.

A clause learning SAT solver would learn the reason for conflict at this 
point and then backtrack to a level such that the learnt clause is UNIT.  
Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
(\(n0: c:=[0,0]\)), that is all error traces must satisfy $(c \neq 1)$ 
at node $n0$.  The analysis discards all interval constraints that
lead to the conflict and backtrack to decision level 0.  ACDCL then 
performs interval analysis with the learnt constraint $(c \neq 1)$ which also 
leads to a conflict, as shown in right side of Figure~\ref{ssa}.  The 
analysis cannot backtrack further and therefore it terminates proving 
that the program is safe.  A DPLL-style solver would perform proof by 
cases.  However, decision and clause learning are used to avoid case 
based reasoning and prevents enumeration behavior.

For $N=46000$, table~\ref{solver} shows the statistics from 
MiniSAT~\cite{minisat} solver for analysis using BMC, and 
analysis using ACDCL with interval and octagon domains.  
Compared to a SAT solver, there is a significant reduction 
in the number of decisions, propagations, learnt clauses and 
restrarts.  Compared to abstract interpretation, ACDCL does 
not require external hints to proof the program, thus 
automatically performing program and property driven trace 
partitioning to generate proofs using decision and clause learning.