### 2016-10-20 conclusion.tex
compile

\section{Conclusions}
In this paper, we present a procedure for lifting the model search 
in satisfiability solvers to relational abstract domains.  We explain 
specific properties of abstract domain elements that are necessary to infer 
domain-specific deductions.  We present decision heuristics that takes 
into account the high-level structure of programs and the underlying 
expressivity of the abstract domains.  Our conflict analysis procedure 
learns abstract transformer following UIP computation over relational 
elements.  We instantiate our algorithms as a program analyzer to determine
safety of C programs.  Experimental evaluation shows that ACDCL is able 
to suitably exploit the expressivity of richer abstract domains as well 
as precision of CDCL architecture.

Generalizations of conflict reasons for relational domains 
and non-relational domain such as trace-based abstrac domains is a 
possible direction of future work.  We believe that domain refinement 
in ACDCL will also help to dynamically adjust the precision of the 
deduction in an effective manner.

\Omit{
Experimental evaluation on different benchmark sets shows that
ACDCL is able to suitably exploit the expressivity of richer
abstract domains as well as precision of CDCL architecture.
This is attributed to the intelligent decision heuristics
which exploits the high-level structure of the problem
combined with stronger deductions and clause learning
mechanism aided by the richer abstract domains.
Future work -- domain refinement, incremental solving, 
invariant inference, intelligent $\gamma$-completeness check. 
Implement dedicated abstract domains to exploit full performance.
generalization for relational domain, instantiate acdl with trace-based domains.
}