### 2016-10-20 experiments.tex
experiment discussion

\section{Experimental Results}
We have implemented the instantiation of ACDCL on relational domains
in 2LS verification tool~\cite{2ls}.  We have used the abstract domains
provided by 2LS (intervals, octagons, polyhedra and equalities) for our
experiments. Although the performance of the abstract domain implementations 
in 2LS is not competitive with that of APRON~\cite{apron} library, but 
the domain implementation in 2LS can handle all C operators 
(including bitwise operations) out of the box and support precise
complementation of meet irreducibles, which are necessary for conflict-driven 
learning. Our tool and benchmarks are available online\footnote{http://www.cprover.org/acdcl/}.

We verified a total of 66 benchmarks in ANSI-C, which are derived from 
(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
cycle accurate software models of hardware circuits automatically 
generated using v2c~\cite{mtk2016}, (c) controller code generated 
from Simulink (d) several hand-crafted benchmarks for equivalence 
checking and bounded loop analysis.  All bounded loops are completely 
unrolled before analysis.  We compare our tool with a state-of-the-art 
SAT-based bounded model checker CBMC~\cite{cbmc} and a commercial 
static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
Astr{\'e}e was configured with default set of domains which includes 
interval, bit-field, congruence, trace partitioning and relational 
domains (octagons, polyhedra, zones, equalities, filter).

\Omit {
To enable precise analysis using Astr{\'e}e, all our benchmarks are 
manually instrumented with partition directives which provides external 
hint to the tool to guide the trace partitioning heuristics.  Usually, 
such high-precision is not needed for static analysis, since it makes 
the analysis very expensive.  Without trace partitioning, the 
analysis using Astr{\'e}e shows high degree of imprecision. 
}
\paragraph {\textbf{Precision and Efficiency of Analysis}}
Figure~\ref{fig:results} presents a comparison of the analysis 
using CBMC and ACDCL.  The cumulative statistics corresponding to 
each phase of the algorithm for all 66 benchmarks is presented in 
Table~\ref{result}.  Experimental result shows a reduction of at least 
two orders of magnitude in the number of decisions, propagations and conflicts 
compared to analyis using CBMC.    
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{result}
\end{table}
%
Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16 benchmarks 
without any restarts.  The solver is restarted in 50 cases to avoid spending 
too much time in branches that does not easily lead to a satisfying assignment or 
stronger clause learning.  In contrast, the analysis using ACDCL solved all
66 benchmarks without any restarts.  This is attributed to intelligent decision 
heuristics which exploits the high-level structure of the program combined 
with stronger deductions and clause learning mechanism aided by the richer 
abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
but the analysis using Astr{\'e}e shows high degree of imprecision.  
Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.

\paragraph {\textbf{Propagation Strategy}}      
Figure~\ref{prop-dec}(a) presents a comparison between {\em forward} propagation 
strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of 
propagation strategy influences the total number of decisions and clause 
learning iterations.  Hence, the propagation strategy has a significant 
influence on the run time which can be visualised in Figure~\ref{prop-dec}.  
Unlike the forward propagation, the multi-way strategy may take several 
iterations to reach the fixed-point, but it significantly reduces the 
total number of decisions and conflicts.  This is attributed to the higher 
precision of the meet irreducibles inferred by the multi-way strategy which 
subsequently aids the decision heuristics to make better decisions. 
\Omit {
We observe that the deductions made during chaotic analysis helps the decision 
heuristics to make better decisions compared to forward analysis.  This is 
attributed to the precision of deductions inferred by the chaotic analysis.         
}

\paragraph {\textbf{Decision Heuristics}} Figure~\ref{prop-dec}(b) presents a 
visualization of different decision heuristics in ACDCL.  Random heuristic
performs the best for most safe benchmarks and all bit-vector category 
benchmarks.  Whereas, the ordered heuristic performs better for programs 
with conditional branches since it prioritizes decisions on meet 
irreducibles that appear in conditional branches over meet irreducibles 
that involve numerical variables.  Whereas, the runtimes for berkmin 
heuristic is slightly greater than the random and ordered heuristic 
for our benchmarks. 
\Omit{
Whereas activity-based heuristics such as Berkmin heuristic which 
works well in propositional cases performs best for benchmarks 
that encountered the maximum number of conflicts to prove safety, 
thus allowing the heuristics to choose the decison variable among the set of learnt clauses.   
}

\paragraph {\textbf{Learning}} Learning has a significant influence on
the runtime of ACDCL.  We compare the UIP-based learning technique with 
an analysis that performs chronoligical backtracking without learning. 
The effect of UIP computation allows ACDCL to backtrack non-chronologically 
and guide the model search with a learnt transformer.  However, the analysis 
without learning exhibits case-enumeration behavior and thus immediately 
runs out of capacity for most benchmarks.

%================  
\input{graphs}
%================