### 2016-10-22 algorithm.tex
moved back from bitbucket to github

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKw{return}{return}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{conflict}
\SetKwData{safe}{safe}
\SetKwData{sat}{sat}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{A program in the form of a set of abstract transformers $\abstransset$.}
\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
$\trail \leftarrow \langle\rangle$ \;
$\reasons \leftarrow []$ \;
$\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return (\unsafe,$\abs(\trail)$)}
  $\decisionvar \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
  $\trail \leftarrow \trail \cdot \decisionvar$ \; 
  $\reasons[|\trail|] \leftarrow \top$ \;
  $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$\;
  \Do{$\mathit{result} = \conflict$} {
    \lIf{$\neg \analyzeconflict_{\confheur}(\abstransset,\trail,\reasons)$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $ACDCL_{\propheur,\decheur,\confheur}(\abstransset)$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The algorithm takes a program \pscmt{make this a set of transformers
  (parametrised by the propagation heuristics)}
$\abstransset$ and propagation, decision, and conflict analysis
heuristics as inputs.  
\rmcmt{Approximation of the concrete transformers in 
$\abstransset$ are typically available in abstract domain in the 
form of strongest-post condition or weakest pre-condition. } 
As internal state, it maintains a propagation
trail $\trail$ and a reason trail \pscmt{map?} $\reasons$.
The propagation trail is an initially empty sequence of meet
irreducibles inferred by the abstract model search phase (deductions
and decisions). The reason trail maps the elements of the
propagation trail to transformers $s\in\abstransset$ that were used to
derive them. The \emph{abstract value} $\abs(\trail)$
corresponding to the propagation trail $\trail$ is the
conjunction of the meet irreducibles on the trail:
$\abs(\trail)=\bigsqcap_{m \in \trail}m$ with
$\abs(\trail)=\top$ if $\trail$ is the empty
sequence $\langle\rangle$.

The procedure $\deduce$ computes a greatest fixed point over the
transformers in $\abstransset$ that refines the abstract value,
similar to the BCP step in SAT solvers.  If the result of $\deduce$
returns \textsf{conflict} ($\bot$), the algorithm terminates with
\textsf{safe}.  Else, the analysis enters into the while($\true$) loop
and makes a new decision by a call to $\decide$ which returns a new
meet irreducible $\decisionvar$.
%
\pscmt{what happens if we cannot make any decision any more?
  -- then we should either have deduced $\bot$ or $\abs(\trail)$ must
  be $\gamma$-complete} 
%
We concatenate $\decisionvar$ to the
trail $\trail$.  The decision $\decisionvar$ therefore refines the current
abstract value $\abs(\trail)$ represented by the trail.
% WE ExPLAIN THAT LATER
%However, $\decisionvar$ must be consistent with the state of the
%solver, i.e. $\abs(\trail\cdot \decisionvar)\neq \bot$.
%
For example, a decision in the interval domain restricts the range of 
intervals for variables.
%
We set the corresponding entry in the reason trail $\reasons$ to $\top$
to mark it as a decision.
% so the analysis jumps under a 
%greatest fixed-point \pscmt{???}. Note that the widening operation in abstract interpretation 
%jumps above a least fixed-point, so decisions can be viewed as 
%{\em dual widening} \pscmt{what do we learn?}.

The procedure $\deduce$ is called next to infer new meet irreducibles
taking into account the current decision.  The model search phase
alternates between the decision and deduction until $\deduce$ returns
either \textsf{sat} or \textsf{conflict}.  In the former case, we have
found an abstract value that represents models of $\formula$, which
are counterexamples to the safety problem, and thus we return
\textsf{unsafe}.
%
% has found an abstract value that represents a set of models of $\formula$ or the
% deduction leads to a \textsf{conflict}.  Checking whether the abstract
% model concretises to a set of models of $\formula$; this
% corresponds to a $\gamma$-completeness~\cite{dhk2013-popl} check in
% abstract interpretation.  If $v$ is $\gamma$-complete, then it cannot
% be refined further.  Thus, the algorithm returns the abstract model
% $v$, which is a set of concrete models, and terminates with
% \textsf{unsafe} or the current abstraction is insufficient to
% determine the satisfiability of $\formula$.

In the latter case the algorithm enters in the $\analyzeconflict$
phase to learn the reason for the conflict.   There can be multiple
incomparable reasons for conflict \pscmt{do you have to explain this here?} -- based on the choice of Unique
Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically chooses one
reason.  A learnt clause must include asserting cuts which guarantees
derivation of new meet irreducibles after backtracking. The clause
learning and backtracking continues as long as the result of deduction
is conflicting ($\bot$), that is, the abstract value does not
abstractly satisfy the formula \pscmt{define}.  If no further
backtrack is possible, then the algorithm terminates and $\formula$ is
\textsf{safe}. Else, the algorithm makes a new decision and the above
process is repeated until $\deduce$ returns \textsf{sat} or the
algorithm backtracks to decision level 0 after a conflict in which
case it returns \textsf{safe}.

%\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\trail$?} The state of a ACDCL solver is a tuple 
%of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
%is a sequence of labelled information of the form $(m,s)$ where 
%$m$ is a meet irreducible and $s = \mathsf{decision}$ if $m$ is a decision, 
%or $s = \mathsf{deduction}$ if $m$ is inferred by a deduction. And $S$ is 
%a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 
%stores variable assignments of the form $(p, t)$, where $p$ is 
%a propositional variable that appears at most once in the trail 
%and $t$ is a truth assignment (true or false).  Whereas, a trail 
%$\trail$ in ACDCL contains a sequence of meet irreducibles 
%inferred by deduction or decision phase where a variable in trail 
%can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
%bounds \pscmt{that's specific for a particular domain}.  
%\rmcmt{define trail refinement}
%\Omit {
%A current valuation $v$ is a meet of all elements of trail
%$\trail$, such that $v = \top$ when $\trail$ is 
%empty or $v=\underset{i \geq 0 \wedge i \leq |\trail|}{\meet m_i}$, where 
%$m_i \in \trail$.  A solver is in conflict if some clauses in
%$\abstransset$ is not abstractly satisfied by $v$.
%}
%\pscmt{define ``consistent with trail''}