### 2016-10-22 paper.tex
added dkcmt

\documentclass[a4paper,conference]{llncs}

\usepackage{import}
\usepackage{times}
\usepackage{microtype}
\usepackage{array}
\usepackage{graphicx,wrapfig}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{plotmarks}
\usepackage{pgfplotstable}
\usepackage{filecontents}
\usepackage{pgfplots}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd}
%\usepackage{mathptm}
\usepackage{color}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{alltt}
\usepackage{psfrag}
\usepackage{epsfig}
\usepackage{wasysym} 
%\usepackage{subfigure}
\usepackage{paralist}
\usepackage{dingbat}
\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
\usepackage{hyperref}
\usepackage[subnum]{cases}
\makeatother

\newcommand{\extendedonly}[1]{}
\newcommand{\paperonly}[1]{#1}

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\cbmcv}{\tool{cbmc 5.0}}
\newcommand{\Omit}[1]{}
\newcommand{\mydef}[1]{\begin{definition}#1\end{definition}}
\newcommand{\rmcmt}[1]{{\color{magenta}{#1}}}%#1
\newcommand{\pscmt}[1]{{\color{blue}{#1}}}%#1
\newcommand{\dkcmt}[1]{{\color{red!70!black}{#1}}}%#1

\input{symbols}

\begin{document}

\title{Model Search in Relational Domains}

\author{Rajdeep Mukherjee\inst{1}\hskip 1em  Peter Schrammel\inst{2}\hskip 1em
Daniel Kroening\inst{1}\hskip 1em Tom Melham\inst{1}}

\institute{University of Oxford, UK \and University of Sussex, UK \\
\email{\{rajdeep.mukherjee,kroening,tom.melham\}@cs.ox.ac.uk},
\email{p.schrammel@sussex.ac.uk}}

\maketitle

%===============================================================================
\begin{abstract}
%===============================================================================
%
\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalisation
of the well-known CDCL algorithm that is used in SAT solvers.
%
\Omit {  ACDCL alternates between a model search, which
performs over-approximate deduction with constraint propagation, and a
conflict analysis, which performs under-approximate abduction with heuristic
choice.
}
%
ACDCL has been successfully used with non-relational abstract domains such
as intervals to improve the performance of a floating-point decision
procedure.  Whereas the propositional reasoning used in satisfiability 
solvers can be rather easily lifted from the Boolean lattice to other 
non-relational domains, application to relational domains is 
challenging because of relational properties of deductions.
%
In this paper, we present the non-trivial liftings of the model search and
conflict analysis algorithms for relational abstract domains.  We have
implemented these algorithms inside the 2LS verification tool and
instantiated our analyser over a template polyhedra domain to check safety
properties of C programs.  We~evaluate the performance of our analyser in
comparison with program verifiers that uses propositional-level solvers and
a static analysis tool.
%
\Omit {
on a set of benchmarks drawn from bit-vector regression 
in SV-COMP'16, bit-precise software models of hardware 
circuits auto-generated from v2c tool, and several 
bounds checking benchmarks 
}
We observe a reduction of two orders of magnitude in the number of 
decisions and conflict analyses compared to a satisfiability 
solver and higher precision compared to an abstract interpreter. 
\Omit{
along with stronger deductions and learnt clauses aided by the richer
abstract domains.
}
%
\end{abstract}

%===============================================================================
\input{introduction}
%===============================================================================

%===============================================================================
\input{contributions}
%===============================================================================

%===============================================================================
\input{example}
%===============================================================================

%===============================================================================
\input{domains}
%===============================================================================

%===============================================================================
\input{algorithm}
%===============================================================================

%===============================================================================
\input{deductions}
%===============================================================================

%===============================================================================
\input{decision}
%===============================================================================

%===============================================================================
\input{learning}
%===============================================================================

%===============================================================================
\input{experiments}
%===============================================================================

%===============================================================================
\input{conclusion}
%===============================================================================

%===============================================================================
\input{background}
%===============================================================================

\bibliographystyle{splncs03}
\bibliography{biblio.bib}

\extendedonly{
\input{appendix}
}

\end{document}